---
title: "Atmospheric Correction"
format:
  html:
    toc: false
---

```{r}
#| label: setup
#| include: false
#| message: false
#| warning: false

required_packages <- c("dplyr", "readxl", "leaflet", "htmltools", "stringr")

missing_packages <- required_packages[
  !vapply(required_packages, requireNamespace, logical(1), quietly = TRUE)
]

if (length(missing_packages) > 0) {
  install.packages(missing_packages, repos = "https://cloud.r-project.org")
}

suppressPackageStartupMessages({
  library(dplyr)
  library(readxl)
  library(leaflet)
  library(htmltools)
  library(stringr)
})

station_key <- function(x) {
  str_remove_all(str_to_upper(x), "_")
}

comment_labels <- c(
  zn = "Z. noltei",
  cn = "Cymodocea nodosa",
  sand = "Sand",
  shell = "Shell",
  worm_reef = "Worm reef",
  spartina = "Spartina",
  mangrove = "Mangrove",
  vaucheria = "Vaucheria",
  cyano = "Cyanobacteria",
  green_algae = "Green algae",
  dry_sand = "Dry sand",
  mpb_mud = "MPB mud",
  mbp_mud = "MPB mud",
  algae_mud = "Green algae",
  white_sand = "Sand",
  seagrass = "Z. noltei"
)

normalize_comment_code <- function(x) {
  code <- str_trim(str_to_lower(x))
  code <- str_replace_all(code, "\\s+", "_")

  dplyr::recode(
    code,
    algae_on_mud = "green_algae",
    algae_mud = "green_algae",
    green_algae = "green_algae",
    seagrass = "zn",
    zostera_noltei = "zn",
    zn = "zn",
    mpb = "mpb_mud",
    mbp = "mpb_mud",
    mbp_mud = "mpb_mud",
    mpb_mud = "mpb_mud",
    microphytobenthos_mud = "mpb_mud",
    white_sand = "sand",
    sand = "sand",
    .default = code
  )
}

label_comment <- function(x) {
  labelled <- unname(comment_labels[normalize_comment_code(x)])
  fallback <- str_to_sentence(str_replace_all(x, "_", " "))
  ifelse(is.na(labelled), fallback, labelled)
}

reflectance_avg <- read.table(
  "Data/ASD/202603_Reflectance_avg.txt",
  header = TRUE,
  check.names = FALSE
)

asd_stations <- tibble(
  station = names(reflectance_avg)[-1],
  station_key = station_key(station)
)

asd_log <- read.delim(
  "Data/ASD/202603_Seagrass_cover.txt",
  sep = "\t",
  stringsAsFactors = FALSE,
  check.names = FALSE
)

gps_log <- read_excel("Data/GPS/Info spectral measure_biomass core_BdA 2026.xlsx")

gps_points <- gps_log %>%
  transmute(
    station_key = station_key(station),
    gps_waypoint = suppressWarnings(as.integer(waypoint)),
    acquisition_date = as.Date(date),
    location = location,
    lat = suppressWarnings(as.numeric(lat)),
    lon = suppressWarnings(as.numeric(lon))
  ) %>%
  filter(!is.na(lat), !is.na(lon)) %>%
  distinct(station_key, gps_waypoint, .keep_all = TRUE)

gps_by_waypoint <- gps_points %>%
  filter(!is.na(gps_waypoint)) %>%
  distinct(gps_waypoint, .keep_all = TRUE) %>%
  select(gps_waypoint, acquisition_date, location, lat, lon)

gps_by_station <- gps_points %>%
  filter(station_key != "") %>%
  distinct(station_key, .keep_all = TRUE) %>%
  select(station_key, acquisition_date, location, lat, lon)

mauritania_samples <- asd_log %>%
  transmute(
    station = station,
    station_key = station_key(station),
    gps_waypoint = suppressWarnings(as.integer(gps)),
    target = label_comment(comment)
  ) %>%
  semi_join(asd_stations, by = "station_key") %>%
  left_join(
    gps_by_waypoint %>%
      rename_with(~ paste0(.x, "_waypoint"), c(acquisition_date, location, lat, lon)),
    by = "gps_waypoint"
  ) %>%
  left_join(
    gps_by_station %>%
      rename_with(~ paste0(.x, "_station"), c(acquisition_date, location, lat, lon)),
    by = "station_key"
  ) %>%
  mutate(
    acquisition_date = coalesce(acquisition_date_waypoint, acquisition_date_station),
    location = coalesce(location_waypoint, location_station),
    lat = coalesce(lat_waypoint, lat_station),
    lon = coalesce(lon_waypoint, lon_station)
  ) %>%
  select(station, station_key, gps_waypoint, target, acquisition_date, location, lat, lon) %>%
  filter(
    !is.na(lat),
    !is.na(lon),
    between(lat, 14, 28),
    between(lon, -18, -4)
  ) %>%
  mutate(
    date_group = format(acquisition_date, "%Y-%m-%d"),
    popup = paste0(
      "<strong>", htmlEscape(station), "</strong><br>",
      htmlEscape(target), "<br>",
      htmlEscape(location), "<br>",
      ifelse(is.na(acquisition_date), "NA", format(acquisition_date, "%Y-%m-%d"))
    )
  ) %>%
  arrange(acquisition_date, station)

date_groups <- sort(unique(mauritania_samples$date_group))
target_levels <- sort(unique(mauritania_samples$target))
target_palette <- colorFactor(
  palette = grDevices::hcl.colors(length(target_levels), "Dark 3"),
  domain = target_levels
)
```

During the week of the fieldwork, ASD measurment have been acquired at the different location of the meadow visited by the team (@fig-Map_ASD_samples). These measurment covers a large range of target, from bare sediment to seagrasses and polychaete reefs. On the 27th of March, nowing that a Sentinel-2 acquisition was planned, special measurement over bright and stable targets (White and Dry Sand) were acquired in order to test the accuracy of the atmospheric correction applied to Sentinel-2 imagery (Sen2Cor).

```{r}
#| fig-cap: Map of ASD sampling stations in Mauritania. Only stations with valid GPS coordinates are shown.
#| label: fig-Map_ASD_samples
#| echo: false
#| error: false
#| message: false
#| eval: true
#| warning: false
#| out-width: "95%"

leaflet(mauritania_samples, options = leafletOptions(preferCanvas = TRUE)) %>%
  addProviderTiles(providers$Esri.WorldImagery, group = "Satellite") %>%
  addProviderTiles(providers$OpenStreetMap.Mapnik, group = "OpenStreetMap") %>%
  {
    map <- .
    for (date_value in date_groups) {
      date_data <- mauritania_samples %>%
        filter(date_group == date_value)

      map <- map %>%
        addCircleMarkers(
          data = date_data,
          lng = ~lon,
          lat = ~lat,
          radius = 7,
          color = ~target_palette(target),
          fillColor = ~target_palette(target),
          fillOpacity = 0.9,
          stroke = TRUE,
          weight = 1,
          popup = ~popup,
          label = ~station,
          group = date_value
        )
    }
    map
  } %>%
  addLayersControl(
    baseGroups = c("Satellite", "OpenStreetMap"),
    overlayGroups = date_groups,
    options = layersControlOptions(collapsed = FALSE)
  ) %>%
  addLegend(
    position = "bottomright",
    pal = target_palette,
    values = mauritania_samples$target,
    title = "Target",
    opacity = 0.95
  ) %>%
  addScaleBar(position = "bottomleft") %>%
  fitBounds(
    lng1 = min(mauritania_samples$lon),
    lat1 = min(mauritania_samples$lat),
    lng2 = max(mauritania_samples$lon),
    lat2 = max(mauritania_samples$lat)
  )
```
