---
title: "Atmospheric Correction"
format:
  html:
    toc: false
editor_options: 
  chunk_output_type: console
---

```{r}
#| label: setup
#| include: false
#| message: false
#| warning: false

required_packages <- c(
  "dplyr",
  "ggiraph",
  "ggplot2",
  "htmltools",
  "leaflet",
  "patchwork",
  "readxl",
  "stringr",
  "tidyr"
)

missing_packages <- required_packages[
  !vapply(required_packages, requireNamespace, logical(1), quietly = TRUE)
]

if (length(missing_packages) > 0) {
  install.packages(missing_packages, repos = "https://cloud.r-project.org")
}

suppressPackageStartupMessages({
  library(dplyr)
  library(ggiraph)
  library(ggplot2)
  library(readxl)
  library(leaflet)
  library(htmltools)
  library(stringr)
  library(tidyr)
})

station_key <- function(x) {
  str_remove_all(str_to_upper(x), "_")
}

comment_labels <- c(
  zn = "Z. noltei",
  cn = "Cymodocea nodosa",
  sand = "Sand",
  shell = "Shell",
  worm_reef = "Worm reef",
  spartina = "Spartina",
  mangrove = "Mangrove",
  vaucheria = "Vaucheria",
  cyano = "Cyanobacteria",
  green_algae = "Green algae",
  dry_sand = "Dry sand",
  mpb_mud = "MPB mud",
  mbp_mud = "MPB mud",
  algae_mud = "Green algae",
  white_sand = "Sand",
  seagrass = "Z. noltei"
)

normalize_comment_code <- function(x) {
  code <- str_trim(str_to_lower(x))
  code <- str_replace_all(code, "\\s+", "_")

  dplyr::recode(
    code,
    algae_on_mud = "green_algae",
    algae_mud = "green_algae",
    green_algae = "green_algae",
    seagrass = "zn",
    zostera_noltei = "zn",
    zn = "zn",
    mpb = "mpb_mud",
    mbp = "mpb_mud",
    mbp_mud = "mpb_mud",
    mpb_mud = "mpb_mud",
    microphytobenthos_mud = "mpb_mud",
    white_sand = "sand",
    sand = "sand",
    .default = code
  )
}

label_comment <- function(x) {
  labelled <- unname(comment_labels[normalize_comment_code(x)])
  fallback <- str_to_sentence(str_replace_all(x, "_", " "))
  ifelse(is.na(labelled), fallback, labelled)
}

reflectance_avg <- read.table(
  "Data/ASD/202603_Reflectance_avg.txt",
  header = TRUE,
  check.names = FALSE
)

srf_path <- "Data/COPE-GSEG-EOPG-TN-15-0007 - Sentinel-2 Spectral Response Functions 2024 - 4.0.xlsx"
s2a_srf <- read_excel(srf_path, sheet = "Spectral Responses (S2A)")
s2a_eq_wavelengths <- read_excel(
  srf_path,
  sheet = "Equivalent Wavelengths",
  skip = 1,
  col_names = c("band", "S2A", "S2B", "S2C", "blank1", "blank2", "definition")
) %>%
  filter(str_detect(band, "^B")) %>%
  transmute(
    band = str_replace(band, "^B0", "B"),
    band_center = as.numeric(S2A)
  )

asd_stations <- tibble(
  station = names(reflectance_avg)[-1],
  station_key = station_key(station)
)

asd_log <- read.delim(
  "Data/ASD/202603_Seagrass_cover.txt",
  sep = "\t",
  stringsAsFactors = FALSE,
  check.names = FALSE
)

gps_log <- read_excel("Data/GPS/Info spectral measure_biomass core_BdA 2026.xlsx")

gps_points <- gps_log %>%
  transmute(
    station_key = station_key(station),
    gps_waypoint = suppressWarnings(as.integer(waypoint)),
    acquisition_date = as.Date(date),
    location = location,
    lat = suppressWarnings(as.numeric(lat)),
    lon = suppressWarnings(as.numeric(lon))
  ) %>%
  filter(!is.na(lat), !is.na(lon)) %>%
  distinct(station_key, gps_waypoint, .keep_all = TRUE)

gps_by_waypoint <- gps_points %>%
  filter(!is.na(gps_waypoint)) %>%
  distinct(gps_waypoint, .keep_all = TRUE) %>%
  select(gps_waypoint, acquisition_date, location, lat, lon)

gps_by_station <- gps_points %>%
  filter(station_key != "") %>%
  distinct(station_key, .keep_all = TRUE) %>%
  select(station_key, acquisition_date, location, lat, lon)

mauritania_samples <- asd_log %>%
  transmute(
    station = station,
    station_key = station_key(station),
    gps_waypoint = suppressWarnings(as.integer(gps)),
    target = label_comment(comment)
  ) %>%
  semi_join(asd_stations, by = "station_key") %>%
  left_join(
    gps_by_waypoint %>%
      rename_with(~ paste0(.x, "_waypoint"), c(acquisition_date, location, lat, lon)),
    by = "gps_waypoint"
  ) %>%
  left_join(
    gps_by_station %>%
      rename_with(~ paste0(.x, "_station"), c(acquisition_date, location, lat, lon)),
    by = "station_key"
  ) %>%
  mutate(
    acquisition_date = coalesce(acquisition_date_waypoint, acquisition_date_station),
    location = coalesce(location_waypoint, location_station),
    lat = coalesce(lat_waypoint, lat_station),
    lon = coalesce(lon_waypoint, lon_station)
  ) %>%
  select(station, station_key, gps_waypoint, target, acquisition_date, location, lat, lon) %>%
  filter(
    !is.na(lat),
    !is.na(lon),
    between(lat, 14, 28),
    between(lon, -18, -4)
  ) %>%
  mutate(
    map_target = ifelse(target == "Dry sand", "Sand", target),
    date_group = format(acquisition_date, "%Y-%m-%d"),
    popup = paste0(
      "<strong>", htmlEscape(station), "</strong><br>",
      htmlEscape(map_target), "<br>",
      htmlEscape(location), "<br>",
      ifelse(is.na(acquisition_date), "NA", format(acquisition_date, "%Y-%m-%d"))
    )
  ) %>%
  arrange(acquisition_date, station)

date_groups <- sort(unique(mauritania_samples$date_group))
target_levels <- sort(unique(mauritania_samples$target))
target_colors <- setNames(
  grDevices::hcl.colors(length(target_levels), "Dark 3"),
  target_levels
)
target_palette <- colorFactor(
  palette = target_colors,
  domain = target_levels
)
map_target_levels <- sort(unique(mauritania_samples$map_target))
map_target_colors <- setNames(target_colors[map_target_levels], map_target_levels)
map_target_palette <- colorFactor(
  palette = map_target_colors,
  domain = map_target_levels
)

ac_station_lookup <- asd_log %>%
  transmute(
    station = station,
    station_key = station_key(station),
    target = label_comment(comment)
  ) %>%
  filter(str_detect(station, "^AC0[1-8]$")) %>%
  distinct(station_key, .keep_all = TRUE)

ac_spectra_hyper <- reflectance_avg %>%
  select(wl, matches("^AC_", ignore.case = TRUE)) %>%
  pivot_longer(
    cols = -wl,
    names_to = "station_column",
    values_to = "reflectance"
  ) %>%
  mutate(station_key = station_key(station_column)) %>%
  left_join(ac_station_lookup, by = "station_key") %>%
  filter(!is.na(station)) %>%
  mutate(
    station = factor(station, levels = paste0("AC0", 1:8)),
    tooltip = paste0(
      as.character(station),
      "<br>",
      target,
      "<br>",
      wl,
      " nm",
      "<br>Reflectance: ",
      signif(reflectance, 4)
    )
  )

asd_wl_min <- min(ac_spectra_hyper$wl, na.rm = TRUE)
asd_wl_max <- max(ac_spectra_hyper$wl, na.rm = TRUE)

s2a_srf_long <- s2a_srf %>%
  rename(wl = SR_WL) %>%
  pivot_longer(
    cols = -wl,
    names_to = "band_column",
    values_to = "srf"
  ) %>%
  mutate(
    band = str_remove(band_column, "^S2A_SR_AV_")
  )

usable_s2a_bands <- s2a_srf_long %>%
  group_by(band) %>%
  summarise(
    overlap_weight = sum(
      ifelse(dplyr::between(wl, asd_wl_min, asd_wl_max), srf, 0),
      na.rm = TRUE
    ),
    .groups = "drop"
  ) %>%
  filter(overlap_weight > 0) %>%
  inner_join(s2a_eq_wavelengths, by = "band") %>%
  arrange(band_center)

ac_spectra_resampled <- ac_spectra_hyper %>%
  select(station, target, wl, reflectance) %>%
  inner_join(
    s2a_srf_long %>%
      inner_join(
        usable_s2a_bands %>% select(band, band_center),
        by = "band"
      ) %>%
      filter(dplyr::between(wl, asd_wl_min, asd_wl_max)),
    by = "wl",
    relationship = "many-to-many"
  ) %>%
  group_by(station, target, band, band_center) %>%
  summarise(
    reflectance = sum(reflectance * srf, na.rm = TRUE) / sum(srf, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    station = factor(station, levels = paste0("AC0", 1:8)),
    band = factor(band, levels = usable_s2a_bands$band),
    tooltip = paste0(
      as.character(station),
      "<br>",
      target,
      "<br>",
      as.character(band),
      " (",
      round(band_center, 1),
      " nm)",
      "<br>Reflectance: ",
      signif(reflectance, 4)
    )
  )
```

Field ASD spectra were collected across the Banc d'Arguin intertidal area to sample the dominant benthic targets, including bare sediments, seagrasses, algae, and biogenic reef structures (@fig-Map_ASD_samples). 

```{r}
#| fig-cap: Map of ASD sampling stations in Mauritania. Only stations with valid GPS coordinates are shown.
#| label: fig-Map_ASD_samples
#| echo: false
#| error: false
#| message: false
#| eval: true
#| warning: false
#| out-width: "95%"

leaflet(mauritania_samples, options = leafletOptions(preferCanvas = TRUE)) %>%
  addProviderTiles(providers$Esri.WorldImagery) %>%
  {
    map <- .
    for (date_value in date_groups) {
      date_data <- mauritania_samples %>%
        filter(date_group == date_value)

      map <- map %>%
        addCircleMarkers(
          data = date_data,
          lng = ~lon,
          lat = ~lat,
          radius = 7,
          color = ~map_target_palette(map_target),
          fillColor = ~map_target_palette(map_target),
          fillOpacity = 0.9,
          stroke = TRUE,
          weight = 1,
          popup = ~popup,
          label = ~station,
          group = date_value
        )
    }
    map
  } %>%
  addLayersControl(
    overlayGroups = date_groups,
    options = layersControlOptions(collapsed = FALSE)
  ) %>%
  addLegend(
    position = "bottomright",
    pal = map_target_palette,
    values = mauritania_samples$map_target,
    title = "Target",
    opacity = 0.95
  ) %>%
  addScaleBar(position = "bottomleft") %>%
  fitBounds(
    lng1 = min(mauritania_samples$lon),
    lat1 = min(mauritania_samples$lat),
    lng2 = max(mauritania_samples$lon),
    lat2 = max(mauritania_samples$lat)
  )
```

The atmospheric-correction assessment focuses on the AC stations measured on 2026-03-27, coincident with a Sentinel-2 acquisition. Hyperspectral ASD measurements were resampled with Sentinel-2 spectral response functions to provide band-compatible in situ reflectance for comparison with satellite products (@fig-AC_spectra).

```{r}
#| label: fig-AC_spectra
#| fig-cap: "Spectral reflectance of the 8 ASD stations used for atmospheric correction validation. Both the original hyperspectral reflectance and the resampled reflectance at Sentinel-2 wavelengths are shown."
#| echo: false
#| error: false
#| message: false
#| eval: true
#| warning: false
#| out-width: "95%"

hyperspectral_plot <- ggplot(
  ac_spectra_hyper,
  aes(
    x = wl,
    y = reflectance,
    color = target,
    group = station
  )
) +
  geom_line_interactive(
    aes(
      tooltip = tooltip,
      data_id = station
    ),
    linewidth = 0.9,
    alpha = 0.9
  ) +
  scale_color_manual(values = target_colors, drop = FALSE) +
  labs(
    x = "Wavelength (nm)",
    y = "Reflectance",
    color = "Target"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "right",
    panel.grid.minor = element_blank()
  )

hyperspectral_widget <- girafe(
  ggobj = hyperspectral_plot,
  width_svg = 9,
  height_svg = 5
) %>%
  girafe_options(
    opts_hover(css = "stroke-width:2.2px;"),
    opts_hover_inv(css = "opacity:0.15;"),
    opts_toolbar(saveaspng = TRUE)
  )

resampled_plot <- ggplot(
  ac_spectra_resampled,
  aes(
    x = band_center,
    y = reflectance,
    color = target,
    group = station
  )
) +
  geom_line(linewidth = 0.8, alpha = 0.8) +
  geom_point_interactive(
    aes(
      tooltip = tooltip,
      data_id = paste(station, band, sep = "_")
    ),
    size = 2.4
  ) +
  scale_color_manual(values = target_colors, drop = FALSE) +
  scale_x_continuous(
    breaks = usable_s2a_bands$band_center,
    labels = round(usable_s2a_bands$band_center)
  ) +
  labs(
    x = "Wavelength (nm)",
    y = "Reflectance",
    color = "Target"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "right",
    panel.grid.minor = element_blank()
  )

resampled_widget <- girafe(
  ggobj = resampled_plot,
  width_svg = 9,
  height_svg = 5
) %>%
  girafe_options(
    opts_hover(css = "stroke-width:2.2px;"),
    opts_hover_inv(css = "opacity:0.15;"),
    opts_toolbar(saveaspng = TRUE)
  )

tagList(
  tags$div(
    class = "spectra-toggle",
    tags$span("Show:"),
    tags$label(
      tags$input(
        type = "radio",
        name = "ac-spectra-view",
        value = "hyperspectral",
        checked = NA
      ),
      "Hyperspectral reflectance"
    ),
    tags$label(
      tags$input(
        type = "radio",
        name = "ac-spectra-view",
        value = "resampled"
      ),
      "S2A resampled reflectance"
    )
  ),
  tags$div(id = "ac-spectra-hyperspectral", hyperspectral_widget),
  tags$div(
    id = "ac-spectra-resampled",
    style = "display:none;",
    resampled_widget
  ),
  tags$script(HTML(
    "(function () {
      const radios = document.querySelectorAll('input[name=\"ac-spectra-view\"]');
      const hyper = document.getElementById('ac-spectra-hyperspectral');
      const resampled = document.getElementById('ac-spectra-resampled');
      function updateView() {
        const checked = document.querySelector('input[name=\"ac-spectra-view\"]:checked');
        if (!checked) return;
        hyper.style.display = checked.value === 'hyperspectral' ? '' : 'none';
        resampled.style.display = checked.value === 'resampled' ? '' : 'none';
      }
      radios.forEach(function (radio) {
        radio.addEventListener('change', updateView);
      });
      updateView();
    })();"
  ))
)
```

Satellite reflectance was extracted at ASD station coordinates from two atmospheric correction products: the standard Sentinel-2 Level-2A Sen2Cor product and ACOLITE-derived surface reflectance (`rhos`). Matchups were assessed using linear regression alongside error metrics comparing image-derived reflectance with ASD measurements resampled to Sentinel-2 spectral bands.

@fig-S2_matchup presents the matchup results for both atmospheric correction approaches—Sen2Cor (left) and ACOLITE (right)—using imagery acquired on the same day as the field measurements (27 March 2026). Both methods exhibit strong agreement with ASD data, with high coefficients of determination (R² = 0.935 for Sen2Cor and 0.92 for ACOLITE).

For low reflectance values (R < 0.3), Sen2Cor-derived reflectance is relatively close to ASD measurements; however, the divergence increases at higher reflectance levels. In contrast, ACOLITE shows consistent agreement with ASD across the full reflectance range.

Although Sen2Cor yields a slightly higher R², it departs more from the 1:1 relationship. Conversely, ACOLITE remains more closely aligned with the 1:1 line, a pattern also supported by the other error metrics, indicating a better overall agreement with field measurements.

```{r}
#| label: s2-matchup-data
#| include: false
#| message: false
#| warning: false
#| cache: false

if (!requireNamespace("MapRs", quietly = TRUE)) {
  stop("MapRs must be installed before rendering this section.")
}
if (!requireNamespace("terra", quietly = TRUE)) {
  stop("terra must be installed before rendering this section.")
}

library(terra)

ac_points <- mauritania_samples %>%
  filter(station %in% paste0("AC0", 1:8)) %>%
  select(station, target, lon, lat) %>%
  distinct() %>%
  arrange(station)

s2b_srf <- read_excel(srf_path, sheet = "Spectral Responses (S2B)")
s2b_eq_wavelengths <- read_excel(
  srf_path,
  sheet = "Equivalent Wavelengths",
  skip = 1,
  col_names = c("band", "S2A", "S2B", "S2C", "blank1", "blank2", "definition")
) %>%
  filter(str_detect(band, "^B")) %>%
  transmute(
    band = str_replace(band, "^B0", "B"),
    band_center = as.numeric(S2B)
  )

s2b_srf_long <- s2b_srf %>%
  rename(wl = SR_WL) %>%
  pivot_longer(
    cols = -wl,
    names_to = "band_column",
    values_to = "srf"
  ) %>%
  mutate(
    band = str_remove(band_column, "^S2B_SR_AV_")
  )

usable_s2b_bands <- s2b_srf_long %>%
  group_by(band) %>%
  summarise(
    overlap_weight = sum(
      ifelse(dplyr::between(wl, asd_wl_min, asd_wl_max), srf, 0),
      na.rm = TRUE
    ),
    .groups = "drop"
  ) %>%
  filter(overlap_weight > 0) %>%
  inner_join(s2b_eq_wavelengths, by = "band") %>%
  arrange(band_center)

ac_spectra_resampled_s2b <- ac_spectra_hyper %>%
  select(station, target, wl, reflectance) %>%
  inner_join(
    s2b_srf_long %>%
      inner_join(
        usable_s2b_bands %>% select(band, band_center),
        by = "band"
      ) %>%
      filter(dplyr::between(wl, asd_wl_min, asd_wl_max)),
    by = "wl",
    relationship = "many-to-many"
  ) %>%
  group_by(station, target, band, band_center) %>%
  summarise(
    reflectance = sum(reflectance * srf, na.rm = TRUE) / sum(srf, na.rm = TRUE),
    .groups = "drop"
  )

stable_reference_points <- gps_log %>%
  transmute(
    station = station,
    station_key = station_key(station),
    cover = suppressWarnings(as.numeric(cover)),
    comment_code = normalize_comment_code(comment),
    target = "Sand",
    Zone = ifelse(is.na(Zone) | str_trim(Zone) == "", "Unknown", Zone),
    lat = suppressWarnings(as.numeric(lat)),
    lon = suppressWarnings(as.numeric(lon))
  ) %>%
  semi_join(asd_stations, by = "station_key") %>%
  filter(
    !is.na(lat),
    !is.na(lon),
    between(lat, 14, 28),
    between(lon, -18, -4)
  ) %>%
  filter(
    comment_code %in% c("sand", "dry_sand"),
    station_key != "D06_G"
  ) %>%
  distinct(station_key, .keep_all = TRUE) %>%
  select(station, target, Zone, lat, lon) %>%
  arrange(Zone, station)

stable_reference_station_lookup <- stable_reference_points %>%
  mutate(station_key = station_key(station)) %>%
  select(station, station_key, target, Zone)

stable_spectra_resampled_s2b <- reflectance_avg %>%
  pivot_longer(
    cols = -wl,
    names_to = "station_column",
    values_to = "reflectance"
  ) %>%
  mutate(station_key = station_key(station_column)) %>%
  inner_join(stable_reference_station_lookup, by = "station_key") %>%
  select(station, Zone, target, wl, reflectance) %>%
  inner_join(
    s2b_srf_long %>%
      inner_join(
        usable_s2b_bands %>% select(band, band_center),
        by = "band"
      ) %>%
      filter(dplyr::between(wl, asd_wl_min, asd_wl_max)),
    by = "wl",
    relationship = "many-to-many"
  ) %>%
  group_by(station, Zone, target, band, band_center) %>%
  summarise(
    reflectance = sum(reflectance * srf, na.rm = TRUE) / sum(srf, na.rm = TRUE),
    .groups = "drop"
  )

stable_zone_summary <- stable_reference_points %>%
  count(Zone, name = "n_stations") %>%
  arrange(Zone)

stable_zone_summary_text <- paste0(
  stable_zone_summary$Zone,
  " (",
  stable_zone_summary$n_stations,
  ")",
  collapse = ", "
)

format_matchup_metric <- function(x, suffix = "") {
  if (!is.finite(x)) {
    return(paste0("NA", suffix))
  }

  paste0(format(round(x, 3), nsmall = 3, trim = TRUE), suffix)
}

compute_matchup_metrics <- function(df, image_label) {
  difference <- df$s2_reflectance - df$asd_reflectance
  valid_pct <- !is.na(df$asd_reflectance) & df$asd_reflectance != 0 & !is.na(difference)
  panel_min <- min(c(0, df$asd_reflectance, df$s2_reflectance), na.rm = TRUE)
  panel_max <- max(c(df$asd_reflectance, df$s2_reflectance), na.rm = TRUE)
  panel_span <- panel_max - panel_min

  if (!is.finite(panel_span) || panel_span == 0) {
    panel_span <- max(abs(panel_max), 1) * 0.1
  }

  rpd_value <- if (any(valid_pct)) {
    mean(difference[valid_pct] / df$asd_reflectance[valid_pct], na.rm = TRUE) * 100
  } else {
    NA_real_
  }

  apd_value <- if (any(valid_pct)) {
    mean(abs(difference[valid_pct]) / abs(df$asd_reflectance[valid_pct]), na.rm = TRUE) * 100
  } else {
    NA_real_
  }

  r2_value <- summary(lm(s2_reflectance ~ asd_reflectance, data = df))$r.squared
  rmsd_value <- sqrt(mean(difference^2, na.rm = TRUE))
  bias_value <- mean(difference, na.rm = TRUE)
  mad_value <- mean(abs(difference), na.rm = TRUE)

  tibble(
    r2 = r2_value,
    rmsd = rmsd_value,
    bias = bias_value,
    mad = mad_value,
    rpd = rpd_value,
    apd = apd_value,
    panel_min = panel_min,
    panel_max = panel_max,
    x = panel_min + 0.02 * panel_span,
    y = panel_max - 0.02 * panel_span,
    label = paste(
      paste0("S2 image: ", image_label),
      paste0("R² = ", format_matchup_metric(r2_value)),
      paste0("RMSD = ", format_matchup_metric(rmsd_value)),
      paste0("bias = ", format_matchup_metric(bias_value)),
      paste0("MAD = ", format_matchup_metric(mad_value)),
      paste0("RPD = ", format_matchup_metric(rpd_value, " %")),
      paste0("APD = ", format_matchup_metric(apd_value, " %")),
      sep = "\n"
    )
  )
}

build_s2_matchup <- function(
  s2_safe_name,
  image_date,
  points_data,
  asd_resampled_data,
  excluded_stations = NULL,
  facet_var = NULL
) {
  s2_safe_path <- if (dir.exists(s2_safe_name)) {
    s2_safe_name
  } else {
    file.path("Data", "S2", s2_safe_name)
  }

  if (!dir.exists(s2_safe_path)) {
    stop("Sentinel-2 SAFE directory not found: ", s2_safe_path)
  }

  image_date <- as.Date(image_date)
  image_label <- format(image_date, "%Y-%m-%d")

  s2_image <- MapRs::Read_S2(s2_safe_path)

  s2_band_lookup <- tibble(
    raster_band = names(s2_image),
    band = str_replace(names(s2_image), "^B0", "B")
  )

  common_s2_bands <- usable_s2b_bands %>%
    inner_join(s2_band_lookup, by = "band") %>%
    arrange(band_center)

  point_id_cols <- setdiff(names(points_data), c("lat", "lon"))
  points_vect <- terra::vect(points_data, geom = c("lon", "lat"), crs = "EPSG:4326")
  points_vect <- terra::project(points_vect, terra::crs(s2_image))

  s2_extracted <- terra::extract(
    s2_image[[common_s2_bands$raster_band]],
    points_vect
  ) %>%
    as_tibble()

  names(s2_extracted) <- c("ID", str_replace(names(s2_extracted)[-1], "^B0", "B"))

  s2_spectra <- bind_cols(
    points_data %>% select(all_of(point_id_cols)),
    s2_extracted %>% select(-ID)
  ) %>%
    pivot_longer(
      cols = -all_of(point_id_cols),
      names_to = "band",
      values_to = "s2_reflectance"
    ) %>%
    mutate(
      s2_reflectance = (s2_reflectance - 1000) / 10000
    ) %>%
    inner_join(
      common_s2_bands %>% select(band, band_center),
      by = "band"
    )

  asd_matchup <- asd_resampled_data %>%
    rename(asd_reflectance = reflectance)

  join_keys <- intersect(names(asd_matchup), names(s2_spectra))
  join_keys <- setdiff(join_keys, c("asd_reflectance", "s2_reflectance"))

  matchup_data <- asd_matchup %>%
    inner_join(s2_spectra, by = join_keys)

  if (!is.null(excluded_stations)) {
    matchup_data <- matchup_data %>%
      filter(!station %in% excluded_stations)
  }

  matchup_data <- matchup_data %>%
    filter(
      !is.na(asd_reflectance),
      !is.na(s2_reflectance)
    ) %>%
    mutate(
      band = factor(band, levels = common_s2_bands$band),
      image_date = image_date
    )

  if ("Zone" %in% names(matchup_data)) {
    matchup_data <- matchup_data %>%
      mutate(
        tooltip = paste0(
          as.character(station),
          "<br>",
          target,
          "<br>Zone: ",
          Zone,
          "<br>",
          band,
          " (",
          round(band_center, 1),
          " nm)",
          "<br>S2 image date: ",
          image_label,
          "<br>ASD: ",
          signif(asd_reflectance, 4),
          "<br>S2: ",
          signif(s2_reflectance, 4)
        )
      )
  } else {
    matchup_data <- matchup_data %>%
      mutate(
        tooltip = paste0(
          as.character(station),
          "<br>",
          target,
          "<br>",
          band,
          " (",
          round(band_center, 1),
          " nm)",
          "<br>S2 image date: ",
          image_label,
          "<br>ASD: ",
          signif(asd_reflectance, 4),
          "<br>S2: ",
          signif(s2_reflectance, 4)
        )
      )
  }

  overall_metrics <- compute_matchup_metrics(matchup_data, image_label)

  facet_stats <- NULL
  if (!is.null(facet_var) && facet_var %in% names(matchup_data)) {
    facet_stats <- matchup_data %>%
      group_by(.data[[facet_var]]) %>%
      group_modify(~ compute_matchup_metrics(.x, image_label)) %>%
      ungroup()
  }

  list(
    data = matchup_data,
    r2 = overall_metrics$r2[[1]],
    min_value = overall_metrics$panel_min[[1]],
    max_value = overall_metrics$panel_max[[1]],
    image_date = image_date,
    image_label = image_label,
    metrics = overall_metrics,
    facet_var = facet_var,
    facet_stats = facet_stats
  )
}

build_acolite_matchup <- function(
  acolite_dir,
  image_date,
  points_data,
  asd_resampled_data,
  excluded_stations = NULL
) {
  if (!dir.exists(acolite_dir)) {
    stop("ACOLITE directory not found: ", acolite_dir)
  }

  image_date <- as.Date(image_date)
  image_label <- format(image_date, "%Y-%m-%d")

  acolite_files <- list.files(
    acolite_dir,
    pattern = "rhos_[0-9.]+\\.tif$",
    full.names = TRUE,
    ignore.case = TRUE
  )

  if (length(acolite_files) == 0) {
    stop("No ACOLITE rhos TIFF files found in: ", acolite_dir)
  }

  acolite_band_lookup <- tibble(
    raster_file = acolite_files,
    wavelength = as.numeric(str_match(basename(acolite_files), "rhos_([0-9.]+)\\.tif$")[, 2])
  ) %>%
    filter(is.finite(wavelength)) %>%
    rowwise() %>%
    mutate(
      nearest_band_index = which.min(abs(usable_s2b_bands$band_center - wavelength)),
      band = usable_s2b_bands$band[[nearest_band_index]],
      band_center = usable_s2b_bands$band_center[[nearest_band_index]],
      band_delta = abs(band_center - wavelength)
    ) %>%
    ungroup() %>%
    filter(band_delta <= 5) %>%
    arrange(band_center) %>%
    distinct(band, .keep_all = TRUE)

  if (nrow(acolite_band_lookup) == 0) {
    stop("No ACOLITE bands could be matched to Sentinel-2B bands in: ", acolite_dir)
  }

  acolite_image <- terra::rast(acolite_band_lookup$raster_file)
  names(acolite_image) <- acolite_band_lookup$band

  point_id_cols <- setdiff(names(points_data), c("lat", "lon"))
  points_vect <- terra::vect(points_data, geom = c("lon", "lat"), crs = "EPSG:4326")
  points_vect <- terra::project(points_vect, terra::crs(acolite_image))

  acolite_extracted <- terra::extract(acolite_image, points_vect) %>%
    as_tibble()

  names(acolite_extracted) <- c("ID", acolite_band_lookup$band)

  acolite_spectra <- bind_cols(
    points_data %>% select(all_of(point_id_cols)),
    acolite_extracted %>% select(-ID)
  ) %>%
    pivot_longer(
      cols = all_of(acolite_band_lookup$band),
      names_to = "band",
      values_to = "s2_reflectance"
    ) %>%
    inner_join(
      acolite_band_lookup %>% select(band, band_center),
      by = "band"
    )

  asd_matchup <- asd_resampled_data %>%
    rename(asd_reflectance = reflectance)

  join_keys <- intersect(names(asd_matchup), names(acolite_spectra))
  join_keys <- setdiff(join_keys, c("asd_reflectance", "s2_reflectance"))

  matchup_data <- asd_matchup %>%
    inner_join(acolite_spectra, by = join_keys)

  if (!is.null(excluded_stations)) {
    matchup_data <- matchup_data %>%
      filter(!station %in% excluded_stations)
  }

  matchup_data <- matchup_data %>%
    filter(
      !is.na(asd_reflectance),
      !is.na(s2_reflectance)
    ) %>%
    mutate(
      band = factor(band, levels = acolite_band_lookup$band),
      image_date = image_date,
      tooltip = paste0(
        as.character(station),
        "<br>",
        target,
        "<br>",
        band,
        " (",
        round(band_center, 1),
        " nm)",
        "<br>S2 image date: ",
        image_label,
        "<br>ASD: ",
        signif(asd_reflectance, 4),
        "<br>ACOLITE: ",
        signif(s2_reflectance, 4)
      )
    )

  overall_metrics <- compute_matchup_metrics(matchup_data, image_label)

  list(
    data = matchup_data,
    r2 = overall_metrics$r2[[1]],
    min_value = overall_metrics$panel_min[[1]],
    max_value = overall_metrics$panel_max[[1]],
    image_date = image_date,
    image_label = image_label,
    metrics = overall_metrics,
    facet_var = NULL,
    facet_stats = NULL
  )
}

combine_correction_matchups <- function(sen2cor_matchup, acolite_matchup) {
  combined_data <- bind_rows(
    sen2cor_matchup$data %>% mutate(correction = "Sen2Cor"),
    acolite_matchup$data %>% mutate(correction = "ACOLITE")
  ) %>%
    mutate(
      correction = factor(correction, levels = c("Sen2Cor", "ACOLITE")),
      tooltip = paste0(
        as.character(station),
        "<br>",
        target,
        "<br>Correction: ",
        correction,
        "<br>",
        band,
        " (",
        round(band_center, 1),
        " nm)",
        "<br>S2 image date: ",
        sen2cor_matchup$image_label,
        "<br>ASD: ",
        signif(asd_reflectance, 4),
        "<br>Satellite: ",
        signif(s2_reflectance, 4)
      )
    )

  panel_min <- min(c(0, combined_data$asd_reflectance, combined_data$s2_reflectance), na.rm = TRUE)
  panel_max <- max(c(combined_data$asd_reflectance, combined_data$s2_reflectance), na.rm = TRUE)
  panel_span <- panel_max - panel_min

  correction_metrics <- combined_data %>%
    group_by(correction) %>%
    group_modify(~ compute_matchup_metrics(.x, sen2cor_matchup$image_label)) %>%
    ungroup() %>%
    mutate(
      x = panel_min + 0.02 * panel_span,
      y = panel_max - 0.02 * panel_span,
      label = str_replace(label, "^S2 image: [^\\n]+\\n", "")
    )

  list(
    data = combined_data,
    image_date = sen2cor_matchup$image_date,
    image_label = sen2cor_matchup$image_label,
    min_value = panel_min,
    max_value = panel_max,
    correction_metrics = correction_metrics
  )
}

summarise_correction_metrics <- function(comparison_object) {
  comparison_object$correction_metrics %>%
    mutate(
      summary = paste0(
        correction,
        ": R² = ",
        vapply(r2, format_matchup_metric, character(1)),
        ", RMSD = ",
        vapply(rmsd, format_matchup_metric, character(1))
      )
    ) %>%
    pull(summary) %>%
    paste(collapse = "; ")
}

make_matchup_widget <- function(matchup_object) {
  build_matchup_plot <- function(plot_data) {
    ggplot(
      plot_data,
      aes(
        x = asd_reflectance,
        y = s2_reflectance,
        color = target,
        shape = band
      )
    ) +
      geom_abline(
        slope = 1,
        intercept = 0,
        linetype = "dashed",
        color = "grey45"
      ) +
      geom_smooth(
        aes(
          x = asd_reflectance,
          y = s2_reflectance,
          group = 1
        ),
        inherit.aes = FALSE,
        method = "lm",
        se = FALSE,
        linewidth = 0.8,
        color = "black"
      ) +
      geom_point_interactive(
        aes(
          tooltip = tooltip,
          data_id = paste(station, band, sep = "_")
        ),
        size = 2.8,
        alpha = 0.9
      ) +
      scale_color_manual(values = target_colors, drop = FALSE) +
      scale_shape_manual(
        values = c(16, 17, 15, 3, 7, 8, 18, 4, 1),
        drop = FALSE
      ) +
      labs(
        x = "ASD reflectance",
        y = "Sentinel-2 reflectance",
        color = "Target",
        shape = "Band"
      ) +
      theme_minimal(base_size = 12) +
      theme(
        legend.position = "right",
        panel.grid.minor = element_blank()
      )
  }

  if (is.null(matchup_object$facet_var)) {
    ggobj <- build_matchup_plot(matchup_object$data) +
      coord_equal(
        xlim = c(matchup_object$min_value, matchup_object$max_value),
        ylim = c(matchup_object$min_value, matchup_object$max_value),
        expand = TRUE
      ) +
      labs(
        subtitle = paste("Sentinel-2 image date:", matchup_object$image_label)
      ) +
      annotate(
        "label",
        x = matchup_object$metrics$x[[1]],
        y = matchup_object$metrics$y[[1]],
        label = matchup_object$metrics$label[[1]],
        hjust = 0,
        vjust = 1,
        label.size = 0.2,
        size = 3,
        lineheight = 0.95
      )

    width_svg <- 7
    height_svg <- 6
  } else {
    facet_name <- matchup_object$facet_var
    facet_values <- matchup_object$facet_stats[[facet_name]]

    zone_plots <- lapply(facet_values, function(facet_value) {
      plot_metrics <- matchup_object$facet_stats %>%
        filter(.data[[facet_name]] == facet_value) %>%
        slice(1)

      plot_data <- matchup_object$data %>%
        filter(.data[[facet_name]] == facet_value)

      build_matchup_plot(plot_data) +
        coord_equal(
          xlim = c(plot_metrics$panel_min[[1]], plot_metrics$panel_max[[1]]),
          ylim = c(plot_metrics$panel_min[[1]], plot_metrics$panel_max[[1]]),
          expand = TRUE
        ) +
        labs(
          title = as.character(facet_value),
          subtitle = paste("Sentinel-2 image date:", matchup_object$image_label)
        ) +
        annotate(
          "label",
          x = plot_metrics$x[[1]],
          y = plot_metrics$y[[1]],
          label = plot_metrics$label[[1]],
          hjust = 0,
          vjust = 1,
          label.size = 0.2,
          size = 2.8,
          lineheight = 0.92
        )
    })

    ggobj <- patchwork::wrap_plots(
      plotlist = zone_plots,
      ncol = 2,
      guides = "collect"
    )

    width_svg <- 11
    height_svg <- 9
  }

  girafe(
    ggobj = ggobj,
    width_svg = width_svg,
    height_svg = height_svg
  ) %>%
    girafe_options(
      opts_hover(css = "stroke:black;stroke-width:1.5px;"),
      opts_hover_inv(css = "opacity:0.2;"),
      opts_toolbar(saveaspng = TRUE)
    )
}

make_correction_comparison_widget <- function(comparison_object) {
  correction_plot <- ggplot(
    comparison_object$data,
    aes(
      x = asd_reflectance,
      y = s2_reflectance,
      color = target,
      shape = band
    )
  ) +
    geom_abline(
      slope = 1,
      intercept = 0,
      linetype = "dashed",
      color = "grey45"
    ) +
    geom_smooth(
      aes(group = 1),
      method = "lm",
      se = FALSE,
      linewidth = 0.8,
      color = "black"
    ) +
    geom_point_interactive(
      aes(
        tooltip = tooltip,
        data_id = paste(correction, station, band, sep = "_")
      ),
      size = 2.8,
      alpha = 0.9
    ) +
    geom_label(
      data = comparison_object$correction_metrics,
      aes(
        x = x,
        y = y,
        label = label
      ),
      inherit.aes = FALSE,
      hjust = 0,
      vjust = 1,
      label.size = 0.2,
      size = 2.8,
      lineheight = 0.92
    ) +
    facet_wrap(~correction, ncol = 2) +
    scale_color_manual(values = target_colors, drop = FALSE) +
    scale_shape_manual(
      values = c(16, 17, 15, 3, 7, 8, 18, 4, 1, 2, 5, 6, 9, 10, 11),
      drop = FALSE
    ) +
    coord_equal(
      xlim = c(comparison_object$min_value, comparison_object$max_value),
      ylim = c(comparison_object$min_value, comparison_object$max_value),
      expand = TRUE
    ) +
    labs(
      x = "ASD reflectance",
      y = "Satellite-derived reflectance",
      color = "Target",
      shape = "Band",
      subtitle = paste("Sentinel-2 image date:", comparison_object$image_label)
    ) +
    theme_minimal(base_size = 12) +
    theme(
      legend.position = "right",
      panel.grid.minor = element_blank()
    )

  girafe(
    ggobj = correction_plot,
    width_svg = 11,
    height_svg = 6
  ) %>%
    girafe_options(
      opts_hover(css = "stroke:black;stroke-width:1.5px;"),
      opts_hover_inv(css = "opacity:0.2;"),
      opts_toolbar(saveaspng = TRUE)
    )
}

build_ac06_vs_ac0102_matchup <- function(matchup_object) {
  comparison_data <- matchup_object$data %>%
    filter(as.character(station) %in% c("AC01", "AC02", "AC06")) %>%
    mutate(
      source_station = as.character(station),
      comparison = ifelse(
        source_station %in% c("AC01", "AC02"),
        "mean(AC01, AC02)",
        "AC06"
      )
    ) %>%
    group_by(comparison, band, band_center, image_date) %>%
    summarise(
      source_stations = paste(sort(unique(source_station)), collapse = ", "),
      asd_reflectance = mean(asd_reflectance, na.rm = TRUE),
      s2_reflectance = mean(s2_reflectance, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    mutate(
      station = comparison,
      target = "Sand",
      tooltip = paste0(
        comparison,
        "<br>Source station(s): ",
        source_stations,
        "<br>",
        band,
        " (",
        round(band_center, 1),
        " nm)",
        "<br>S2 image date: ",
        matchup_object$image_label,
        "<br>ASD: ",
        signif(asd_reflectance, 4),
        "<br>S2: ",
        signif(s2_reflectance, 4)
      )
    ) %>%
    arrange(comparison, band)

  comparison_metrics <- compute_matchup_metrics(comparison_data, matchup_object$image_label)

  list(
    data = comparison_data,
    r2 = comparison_metrics$r2[[1]],
    min_value = comparison_metrics$panel_min[[1]],
    max_value = comparison_metrics$panel_max[[1]],
    image_date = matchup_object$image_date,
    image_label = matchup_object$image_label,
    metrics = comparison_metrics,
    facet_var = NULL,
    facet_stats = NULL
  )
}

make_ac_sand_comparison_widget <- function(matchup_object) {
  comparison_colors <- c(
    "AC06" = "#0072B2",
    "mean(AC01, AC02)" = "#D55E00"
  )

  comparison_plot <- ggplot(
    matchup_object$data,
    aes(
      x = asd_reflectance,
      y = s2_reflectance,
      color = comparison,
      shape = band
    )
  ) +
    geom_abline(
      slope = 1,
      intercept = 0,
      linetype = "dashed",
      color = "grey45"
    ) +
    geom_smooth(
      aes(
        x = asd_reflectance,
        y = s2_reflectance,
        group = 1
      ),
      inherit.aes = FALSE,
      method = "lm",
      se = FALSE,
      linewidth = 0.8,
      color = "black"
    ) +
    geom_point_interactive(
      aes(
        tooltip = tooltip,
        data_id = paste(comparison, band, sep = "_")
      ),
      size = 2.8,
      alpha = 0.9
    ) +
    scale_color_manual(values = comparison_colors, drop = FALSE) +
    scale_shape_manual(
      values = c(16, 17, 15, 3, 7, 8, 18, 4, 1),
      drop = FALSE
    ) +
    coord_equal(
      xlim = c(matchup_object$min_value, matchup_object$max_value),
      ylim = c(matchup_object$min_value, matchup_object$max_value),
      expand = TRUE
    ) +
    labs(
      x = "ASD reflectance",
      y = "Sentinel-2 reflectance",
      color = "Reference",
      shape = "Band",
      subtitle = paste("Sentinel-2 image date:", matchup_object$image_label)
    ) +
    annotate(
      "label",
      x = matchup_object$metrics$x[[1]],
      y = matchup_object$metrics$y[[1]],
      label = matchup_object$metrics$label[[1]],
      hjust = 0,
      vjust = 1,
      label.size = 0.2,
      size = 3,
      lineheight = 0.95
    ) +
    theme_minimal(base_size = 12) +
    theme(
      legend.position = "right",
      panel.grid.minor = element_blank()
    )

  girafe(
    ggobj = comparison_plot,
    width_svg = 7,
    height_svg = 6
  ) %>%
    girafe_options(
      opts_hover(css = "stroke:black;stroke-width:1.5px;"),
      opts_hover_inv(css = "opacity:0.2;"),
      opts_toolbar(saveaspng = TRUE)
    )
}

matchup_20260327 <- build_s2_matchup(
  "S2B_MSIL2A_20260327T113319_N0512_R080_T28QCH_20260327T152639.SAFE",
  "2026-03-27",
  points_data = ac_points,
  asd_resampled_data = ac_spectra_resampled_s2b,
  excluded_stations = "AC03"
)

matchup_20260327_acolite <- build_acolite_matchup(
  "Data/S2/ACOLITE/20260327",
  "2026-03-27",
  points_data = ac_points,
  asd_resampled_data = ac_spectra_resampled_s2b,
  excluded_stations = "AC03"
)

matchup_20260327_corrections <- combine_correction_matchups(
  matchup_20260327,
  matchup_20260327_acolite
)

matchup_20260327_zone <- build_ac06_vs_ac0102_matchup(matchup_20260327)

matchup_20260307_ac_sand <- build_s2_matchup(
  "S2B_MSIL2A_20260307T113319_N0512_R080_T28QCH_20260307T154921.SAFE",
  "2026-03-07",
  points_data = ac_points,
  asd_resampled_data = ac_spectra_resampled_s2b,
  excluded_stations = "AC03"
)

matchup_20260307_acolite <- build_acolite_matchup(
  "Data/S2/ACOLITE/20260307",
  "2026-03-07",
  points_data = ac_points,
  asd_resampled_data = ac_spectra_resampled_s2b,
  excluded_stations = "AC03"
)

matchup_20260307_corrections <- combine_correction_matchups(
  matchup_20260307_ac_sand,
  matchup_20260307_acolite
)

matchup_20260307_ac_sand_comparison <- build_ac06_vs_ac0102_matchup(matchup_20260307_ac_sand)

matchup_20260307 <- build_s2_matchup(
  "S2B_MSIL2A_20260307T113319_N0512_R080_T28QCH_20260307T154921.SAFE",
  "2026-03-07",
  points_data = stable_reference_points,
  asd_resampled_data = stable_spectra_resampled_s2b,
  facet_var = "Zone"
)
```

```{r}
#| fig-cap: Match-up between ASD reflectance resampled to Sentinel-2B bands and Sentinel-2 image reflectance extracted at the field locations for the image acquired on 2026-03-27. Sen2Cor and ACOLITE atmospheric corrections are compared side by side.
#| label: fig-S2_matchup
#| echo: false
#| error: false
#| message: false
#| eval: true
#| warning: false
#| out-width: "95%"

make_correction_comparison_widget(matchup_20260327_corrections)
```

@fig-S2_matchup_20260307_ac presents the comparison between ASD field measurements acquired on 27 March and satellite-derived reflectance from an image acquired 20 days earlier (7 March). While the temporal mismatch likely contributes to the observed differences, the comparison between atmospheric correction methods remains informative.

Sen2Cor exhibits lower dispersion relative to ASD measurements, with a higher coefficient of determination (R² = 0.845) compared to ACOLITE (R² = 0.795). However, ACOLITE shows a closer alignment with the 1:1 relationship and consistently lower error metrics (RMSD, bias, MAD, RPD, and APD).

Overall, despite the lower R², ACOLITE provides more accurate and consistent reflectance estimates, making it a more suitable choice for long-term time series analyses of the Banc d’Arguin based on satellite-derived data.

```{r}
#| fig-cap: Match-up between ASD reflectance resampled to Sentinel-2B bands and Sentinel-2 image reflectance extracted at the field locations for the image acquired on 2026-03-07. Sen2Cor and ACOLITE atmospheric corrections are compared side by side.
#| label: fig-S2_matchup_20260307_ac
#| echo: false
#| error: false
#| message: false
#| eval: true
#| warning: false
#| out-width: "95%"

make_correction_comparison_widget(matchup_20260307_corrections)
```

## Times series analysis of the reflectance divergeance. 

The reflectance of selected targets, assumed to remain stable over time, will be monitored and compared to the reflectance values acquired on 27 March 2026.

